Dependent Object Types

### 2012-09-07 doc/candidacy/slides.tex
cleanup dead-code
rev

\useoutertheme{infolines}
\setbeamertemplate{headline}{} % removes the headline that infolines inserts
% \setbeamertemplate{footline}{
%   \hfill%
%   \usebeamercolor[fg]{page number in head/foot}%
%   \usebeamerfont{page number in head/foot}%
%   \insertpagenumber\,/\,\insertpresentationendpage\kern1em\vskip2pt%
% }
\setbeamertemplate{footline}{
  \hfill%
  \usebeamercolor[fg]{page number in head/foot}%
  \usebeamerfont{page number in head/foot}%
  \insertpagenumber\kern1em\vskip2pt%
}
\setbeamertemplate{navigation symbols}{}

@@

\begin{frame}{SML modules}
\begin{description}[structure]
\item[structure] a module; a collection of types, values and
  (sub)structures packaged as a program unit.
\item[signature] an interface; the type of a structure; description of
  the types, values and (sub)structures of a structure given by their
  kinds, types and interfaces.
\item[functor] a function mapping structures to structures
\item[type-sharing] enables stating that types specified in separate
  modules are actually equal
\end{description}
% example?
\end{frame}

@@

\begin{frame}{ML-Like: Translucent Sums}
\begin{itemize}
\item translucent sums are values representing modules
\item translucent sum: sequence of bindings
\item translucent sum type: sequence of declarations
\item unlike traditional records, later fields can depend on earlier ones
\item any type can be partially or fully determined by a type expression
\item dependent typing since translucent sums are terms that may contain type components
\end{itemize}
\end{frame}

@@

\begin{frame}{ML-Like: Path-Dependent Types}
\begin{itemize}
\item Scala's abstract types  $\approx$ ML's abstract types of signatures; limitations:
\begin{itemize}
\item recursive references
\item bounded quantifications
\end{itemize}
\item translucent sum can be given a more precise type by referring to the name of its type member with a path selection; essential for %example with justification for two items below?
\begin{itemize}
\item breaking the dependencies between (sub)fields
\item propagating typing information
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}{${\nu}Obj$: Syntax}
\begin{itemize}
\item terms in ${\nu}Obj$ denote objects or classes; consist of:
\begin{description}[mixin compositions]
\item[variables] $x$
\item[selections] $t.l$
\item[object creations] \mbox{$\newin{x \larrowc t}{u}$}
\item[class templates] $\cls{x: S \sep \tup d}$
\item[mixin compositions] $t \withsub{S} u$
\end{description}
\item a value is a variable or class template
\item a path is a variable followed by a possibly empty sequence of selections
\item types in ${\nu}Obj$ consist of:
\begin{description}[mixin compositions]
\item[singleton types] $p.\type$
\item[type selections] $T\Tsel L$
\item[record types] $\obj{x \sep \tup D}$ % explain difference between record and class types?
\item[compound types] $T \with U$
\item[class types] $\cls{x: S \sep \tup D}$
\end{description}
\end{itemize}
\end{frame}

>>>>

\useoutertheme{infolines}
\setbeamertemplate{headline}{}
\setbeamertemplate{footline}{
  \hfill
  \usebeamercolor[fg]{page number in head/foot}
  \usebeamerfont{page number in head/foot}
  \insertpagenumber\kern1em\vskip2pt
}
\setbeamertemplate{navigation symbols}{}

@@

\begin{frame}{SML modules}
\begin{description}[type-sharing]
\item[structure] a module\\a collection of types, values and
  (sub)structures packaged as a program unit.
\item[signature] an interface; the type of a structure\\description of
  the types, values and (sub)structures of a structure given by their
  kinds, types and interfaces.
\item[functor] a function mapping structures to structures
\item[type-sharing] enables stating that types specified in separate
  modules are actually equal
\end{description}
% example?
\end{frame}

@@

\begin{frame}{ML-Like: Translucent Sums}
\begin{itemize}
\item translucent sums are values representing modules
\item translucent sum: sequence of bindings
\item translucent sum type: sequence of declarations
\item unlike traditional records, later fields can depend on earlier ones
\item fields can be types or terms
\item any type can be partially or fully determined by a type expression
\item dependent typing since translucent sums are terms that may contain type components
\end{itemize}
\end{frame}

@@

\begin{frame}{ML-Like: Path-Dependent Types}
\begin{itemize}
\item Scala's abstract types  $\approx$ ML's abstract types of signatures; limitations:
\begin{itemize}
\item recursive references
\item bounded quantifications
\end{itemize}
\item translucent sum can be given a more precise type by referring to the name of its type member with a path selection; essential for %example with justification for two items below? not clear without introducing internal names?
\begin{itemize}
\item breaking the dependencies between (sub)fields
\item propagating typing information
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}{${\nu}Obj$: Syntax}
\begin{itemize}
\item terms in ${\nu}Obj$ denote objects or classes; consist of:
\begin{description}[mixin compositions]
\item[variables] $x$
\item[selections] $t.l$
\item[object creations] \mbox{$\newin{x \larrowc t}{u}$}
\item[class templates] $\cls{x: S \sep \tup d}$
\item[mixin compositions] $t_1 \withsub{S} t_2$\\(if omitted, $S = {S_1 \with (S_2 \with
  \obj{x \sep D_1 \uplus D_2})}$)
\end{description}
\item a value is a variable or class template
\item a path is a variable followed by a possibly empty sequence of selections
\item types in ${\nu}Obj$ consist of:
\begin{description}[mixin compositions]
\item[singleton types] $p.\type$
\item[type selections] $T\Tsel L$
\item[record types] $\obj{x \sep \tup D}$ % explain difference between record and class types?
\item[compound types] $T \with U$
\item[class types] $\cls{x: S \sep \tup D}$
\end{description}
\end{itemize}
\end{frame}
